This project seeks to aims latency and hence enhance data throughput for reading and writing on Iceberg tables inside \gls{HopsFS}. The achievement of this objective is tight to a specified list of \glspl{G}, here described. These are also connected to the collection of \glspl{RQ}, delineating a distinct framework of the different project milestones.

\begin{enumerate}
    \item \glspl{G} aimed to answer RQ1: 
        \begin{enumerate}
            \item[G1:] Understand PyIceberg library tools, and identify which are needed to interact with \gls{HopsFS}.
            \item[G2:] Implement interaction between PyIceberg and \gls{HopsFS}.
            \item[G3:] Design the experiments to evaluate the performance difference between the legacy access to Apache Hudi and the PyIceberg library-based access to Apache Iceberg, on \gls{HopsFS}. 
            \item[G4:] Perform the designed experiments.
            \item[G5:] Visualize the experiments' results, focusing on an effective comparison of performances.
            \item[G6:] Examine and describe the findings in a dedicated Section of the thesis report.
        \end{enumerate}
    \item \glspl{G} aimed to answer RQ2:
        \begin{enumerate}
            \item[G7:] Investigate related work on delta-rs library-based access to Delta Lake, on \gls{HopsFS}.
            \item[G8:] Visualize experiments' results of PyIceberg and delta-rs alternatives to the legacy system, focusing on an effective comparison of performances.
            \item[G9:] Examine and describe the findings in a dedicated Section of the thesis report.
        \end{enumerate}
\end{enumerate}

To reach these \glspl{G}, several \glspl{D} will be created:
\begin{enumerate}
    \item[D1:] Experiment results on the difference in performance between the legacy access to Apache Hudi and the PyIceberg library-based access to Apache Iceberg, on \gls{HopsFS}
    This \gls{D} is related to completing \glspl{G}3--5.
    \item[D2:] Experiment results on the difference in performance between PyIceberg library-based access to Apache Iceberg and the delta-rs library-based access to Delta Lake, on \gls{HopsFS}.
    This \gls{D} is related to completing \gls{G}7 and \gls{G}8.
    \item[D3:] This thesis report. It offers additional detail on the implementation, design choices, performance, and evaluation of the results.
    This \gls{D} a comprehensive report of all thesis work, incorporating the analysis specified in \gls{G}6 and \gls{G}9.
\end{enumerate}