\markboth{Acknowledgments}{}

\begin{comment}
    When you write your acknowledgements, write an exhaustive list of all the people you wish to
    thank for helping or collaborating with you on your thesis; then organize them, beginning with
    those who helped you with the product (the actual writing of the dissertation itself) the most.
    You may even want to say some words about the people who helped you in finding the
    sources for your thesis: the librarian’s knowledge and patience cannot be overestimated. Do
    not overlook those (perhaps before you came to Georgia Tech) who helped you to gain
    practical experience. These people may have been paramount in their contribution to what
    you achieved at Tech. At the end, most writers thank those who lent moral/emotional
    support. Of course, they are often parents, spouses, friends, and colleagues—all of those who
    supported your effort.
\end{comment}

\begin{quote}
    “ Thanks to doctor Heimlich for your manouver...„

    -- Brothers xxx, yyy
\end{quote}
Ever since I first read this sentence in their programming book's acknoledgments, I realized that every challenge, no matter how serious or difficult, can be approached with a big smile. Yet, beyond a positive mindset, the completion of this master's thesis was only possible thanks to the invaluable support of many, to whom I am deeply grateful.

Many thanks to Jim Dowling, how offered me the opportunity to work on this project and Davit Bzhalava, who supported me along the way. I also had the great pleasure of collaborating with the employees of Hopsworks AB, who generously shared their patience and expertise.

Many thanks to my KTH examiner, Vladimir Vlassov, and my supervisors, Fabian Schmidt and Sina Sheikholeslami, for their guidance in conducting thorough research and for their invaluable feedback.

Heartfelt thanks to my parents and relatives, who have supported me, who have inspired me, who have believed in me, giving me a blank canvas with colors they had never used before. Your wisdom, guidance, and the values you instilled in me have shaped who I am today -- and the person I strive to become in the future.

Lastly, I cannot begin to express my gratitude to the friends who have stood by my side through adventures and challenges. Michele, Daniele, and Davide, for countless years, you offered me advice and challenged my ideas, guiding me towards a clearer understanding of myself. Alfonso, Giacomo, Luca, and Virginia, you made feel like home, at every step of this journey, no matter where we were. Giovanni, you have filled my mind with 'whys' and 'hows' -- questions that helped me grow at every answer, though some still await resolution. Giulia, and all the incredible people from MESA, you have shown me how essential it is to have a cause to believe in and care about. May the future bring even more time to grow, learn, and smile -- together.